### 2025-05-08 project3/draft.tex
new

In the evolving landscape of financial markets, insurance products such as \gls{va}s have gained a substantial amount of interest due to their ability to provide both investment growth and guaranteed benefits. 
Managing risks associated with these products in volatile market conditions is a complex task that demands sophisticated financial modeling techniques such as nested simulation.
A standard nested simulation procedure involves running a large number of simulations to generate a dataset of scenario-wise contract losses with two levels of simulations.
Practitioners can estimate tail risk measures using the realized losses for different risk scenarios obtained from these simulations.
Nested simulation is extremely computationally expensive.
To address the computational burden, metamodeling techniques have been proposed, where a surrogate model approximates the outcomes of a complex simulation procedure. 
Traditional \gls{ml} models often struggle to capture the intricate nonlinear relationships and temporal dependencies inherent in financial data.
In particular, \gls{dnn}s, and specifically \gls{lstm} networks, have been employed as metamodels to predict the outcomes of the inner simulations efficiently.
Chapter~\ref{chap:project2} introduced a metamodeling-based nested simulation framework for dynamic hedging of \gls{va}s that approximates the inner simulation by an \gls{lstm} metamodel.
It shows that \gls{lstm}s, as a type of gated \gls{rnn}s, were well-suited as metamodels for \gls{mc} simulation of financial time series and could capture the long-term dependencies in the simulation dataset.
This property is crucial for the application of dynamic hedging of \gls{va}s.

@@

Constraints on the simulation budget is the most critical challenge for dynamic hedging of \gls{va}s.
Any method that involves nested simulation is computationally expensive, and the computational cost is often prohibitive for changing market conditions.
Therefore, it is essential to develop methods that can rapidly adapt existing metamodels to new conditions without incurring the full computational cost of retraining and extensive addtional simulations.
Transfer Learning (\gls{tl}) offers a compelling solution to this problem by enabling the reuse of a pre-trained model on a new but related task~\citep{pan2009survey}.
In the context of \gls{dnn}s, \gls{tl} involves leveraging the knowledge acquired during training on one dataset to improve learning performance on a different dataset~\citep{yosinski2014transferable}.
This approach can substantially reduce the amount of training data needed, training time and computational resources while enhancing model generalization.
Instead of starting from the scratch, a new \gls{dnn} metamodel can be built on top of a pre-trained metamodel and fine-tuned on new data.
This allows the metamodel to adapt quickly to unseen conditions and new \gls{va} contracts.

In this chapter, we explore the application of \gls{tl} to the dynamic hedging of \gls{va}s using \gls{lstm} metamodels.
We propose a novel \gls{tl} framework that accelerates the training of \gls{dnn} metamodels for nested simulation procedures in dynamic hedging.
This setting is particularly relevant for dynamic hedging problems, as insurance companies often underwrite new contracts under changing market conditions and actuarial assumptions.
This chapter proposes a \gls{tl} framework that makes fast adaptation of metamodels possible.
Our approach involves pre-training an \gls{lstm} network on a large dataset of \gls{va} simulations and then fine-tuning it on a smaller dataset of new \gls{va} contracts.
We evaluate the performance of the \gls{tl} framework on a real-world dataset of \gls{va} contracts and compare it with training a \gls{lstm} metamodel from scratch.
More importantly, we show the strengths and weaknesses of different \gls{tl} techniques for rapid adaptation of \gls{lstm} metamodels.

The rest of this chapter is organized as follows.
Section~\ref{sec3:background} provides an overview of the dynamic hedging problem for \gls{va}s and the use of \gls{lstm} networks as metamodels in a nested simulation procedure.
Section~\ref{sec3:transfer_learning} introduces the \gls{tl} framework for rapid adaptation of \gls{lstm} metamodels in dynamic hedging.
Section~\ref{sec3:experiments} presents the experimental setup and results, comparing the performance of \gls{tl} with training from scratch.
Finally, Section~\ref{sec3:conclusion} concludes the chapter and discusses future research directions.

@@

\gls{tl} is a machine learning paradigm where knowledge acquired from a source task is utilized to improve learning performance on a related target task.
One of the primary applications of \gls{tl} in finance is asset price prediction. 
Traditional models, such as geometric Brownian motion (\gls{gbm}), autoregressive integrated moving average (\gls{arima}) and generalized autoregressive conditional heteroskedasticity (\gls{garch}), have been widely used in time series modeling.
\gls{arima} models~\citep{box1970distribution} combine autoregression, differencing, and moving average components to capture linear relationships in time series data.
\gls{garch} models~\citep{bollerslev1990modelling}, extend \gls{arch} models to better capture volatility clustering, which is a key stylized fact of financial time series~\citep{cont2001empirical}.
However, these models often struggle to capture complex dynamics, e.g., regime changes~\citep{hamilton1989new} and extreme market events~\citep{embrechts2013modelling} that are prevalent in financial markets.
In Chapter~\ref{chap:project2}, our numerical experiment uses a regime-switching geometric Brownian motion (\gls{rsgbm}) as the underlying asset model.
For estimating tail risk measures, a metamodel-based nested simulation procedure is proposed to approximate the inner simulation.
We have shown \gls{dnn} metamodels like \gls{rnn} and \gls{lstm} networks have demonstrated substantial improvements in modeling temporal dependencies, and they save a substantial amount of simulation budget.
However, training these models from the scratch requires an extensive amount of simulation data, which may not always be available for specific assets or under certain market conditions.

@@

Regarding a nested simulation of \gls{va}s,~\cite{cheng2019fast} is the most relevant study to our work, where they proposed a \gls{tl} framework for fast valuation of large portfolios of \gls{va}s.
Instead of using stochastic kriging~\citep{gan2015valuation}, they employed a pre-trained \gls{dnn} to select the best representative scenarios from a large portfolio of \gls{va}s.
In our work, we focus on the application of \gls{tl} to accelerate the training of \gls{lstm} metamodels for a nested simulation in dynamic hedging of \gls{va}s.
The fine-tuned \gls{lstm} metamodels can be readily adapted to the two-stage procedure and the single-stage procedure in Chapter~\ref{chap:project2} to predict the contract losses under different scenarios.

In our context of \gls{dnn}s metamodeling-based simulation procedures for hedging \gls{va}s, \gls{tl} involves pre-training a \gls{dnn} metamodel where the simulation budget is abundant and then fine-tuning it on a smaller dataset of new contracts or market conditions.
Written on the same underlying asset, different \gls{va} contracts may share common features or exhibit similar patterns, especially temporal dependencies and regime changes in the underlying financial time series.
Similarly, two \gls{va}s with the same features but based on different underlying assets may have some shared characteristics that can be leveraged to improve the learning performance of the metamodel.
By transferring knowledge from a pre-trained model to a new but related task, \gls{tl} can substantially reduce the computational cost of training the metamodel on the target task.

@@

In this section, we propose a \gls{tl} framework for rapid adaptation of \gls{lstm} metamodels in dynamic hedging of \gls{va}s.
The goal is to leverage the knowledge acquired during training on a large dataset of \gls{va} simulations to improve the learning performance on a smaller dataset of new \gls{va} contracts.

@@

        \STATE \textbf{Multi-task learning:} define shared LSTM layers $\theta_0$ and task-specific fully connected layers $\theta_k$ for each task $k$. The LSTM layers are shared across all tasks $\{\mathcal{T}_k\}_{k=1}^K$.

@@

        \STATE \textbf{Layer freezing:} once the multi-task metamodel is trained, freeze the parameters of the shared LSTM layers $\theta_0$.

@@

In this section, we evaluate the performance of the \gls{tl} framework for rapid adaptation of \gls{lstm} metamodels in dynamic hedging of \gls{va}s.
The low noise dataset generated by the standard nested simulation procedure in Section~\ref{sec2:numerical} is used to train the source \gls{lstm} metamodels.
We consider \gls{tl} to two types of target tasks: 
\begin{itemize}
    \item the same contract but different underlying asset model, and
    \item different contracts but the same asset model.
\end{itemize}  
Similar to Section~\ref{sec2:numerical}, we use the standard nested simulation procedure to generate the dataset for the source and target tasks.
Serveral \gls{va} datasets generated with the standard nested simulation procedures under geometric Brownian motion (\gls{gbm}) and regime-switching \gls{gbm} (\gls{rsgbm}) asset models.
In this experiment, we use the same \gls{rsgbm} asset model as Chapter~\ref{chap:project2} to generate the dataset for the source and target tasks.
When following a static lapse model and a dynamic lapse model, the monthly lapse rate is given by Equation~\ref{eq:lapse_rate} and Equation~\ref{eq:dynamic_lapse_rate}, respectively.
The experiments are run on a machine with a AMD Ryzen 9 7900X processor with 32 GB of RAM and a Nvidia RTX $\num{3060}$ Ti GPU.

@@

Multi-task learning enables the \gls{lstm} metamodels to learn shared representations across related \gls{va} contracts.
In this section, we examine the performance of multi-task learning applied to two types of \gls{va}s, \gls{gmmb} and \gls{gmwb} with the dynamic lapse rates.
The simulation datasets contain $2,\!000$ samples for each contract type, and the \gls{lstm} metamodels are trained using multi-task learning. 
In our experiments, Algorithm~\ref{alg3:multiTaskLearning} is used to train the \gls{lstm} metamodels simultaneously to minimize the multi-task MSE loss function in Equation~\eqref{eq3:multiTaskLoss}.
We use individual task training as a baseline for comparison, where the \gls{lstm} metamodels are trained separately on the \gls{gmmb} and \gls{gmwb} contracts.
The objective is to assess how multi-task learning can improve the training efficiency and performance of both products compared to individual task training.

@@

Multi-task learning offers a robust framework for transferring knowledge across related simulation schemes in the \gls{lstm} metamodeling for nested simulations. 
By sharing representations through the \gls{lstm} layers, the model can learn more generalizable features that are beneficial across different \gls{va} contracts. 
The multi-task approach is particularly beneficial when training data for individual tasks is scarce, as it effectively pools information across tasks to enhance learning efficiency and predictive performance.
We have extended the idea of pooling in Chapter~\ref{chap:project1} and~\ref{chap:project2} to a higher level.
Furthermore, a metamodel trained with multi-task learning can serve as a pre-trained model. 
It is adaptable to various \gls{va} contracts and asset models, and it is a versatile tool for practical applications in dynamic hedging of \gls{va}s.

>>>>

In the evolving landscape of financial markets, insurance products such as \gls{va} contracts have gained a substantial amount of interest due to their ability to provide both investment growth and guaranteed benefits. 
Managing risks associated with these products in volatile market conditions is a complex task that demands sophisticated financial modeling techniques such as nested simulation.
A standard nested simulation procedure involves running a large number of simulations to generate a dataset of scenario-wise contract losses with two levels of simulations.
Practitioners can estimate tail risk measures using the realized losses for different risk scenarios obtained from these simulations.
Nested simulation is extremely computationally expensive.
To address the computational burden, metamodeling techniques have been proposed, where a surrogate model approximates the outcomes of a complex simulation procedure. 
Traditional \gls{ml} models often struggle to capture the intricate nonlinear relationships and temporal dependencies inherent in financial data.
In particular, \gls{dnn}s, and specifically \gls{lstm} networks, have been employed as metamodels to predict the outcomes of the inner simulations efficiently.
Chapter~\ref{chap:project2} introduced a metamodeling-based nested simulation framework for dynamic hedging of \gls{va} contracts that approximates the inner simulation by an \gls{lstm} metamodel.
It shows that \gls{lstm}s, as a type of gated \gls{rnn}s, were well-suited as metamodels for \gls{mc} simulation of financial time series and could capture the long-term dependencies in the simulation dataset.
This property is crucial for the application of dynamic hedging of \gls{va} contracts.

@@

Constraints on the simulation budget is the most critical challenge for dynamic hedging of \gls{va} contracts.
Any method that involves nested simulation is computationally expensive, and the computational cost is often prohibitive for changing market conditions.
Therefore, it is essential to develop methods that can rapidly adapt existing metamodels to new conditions without incurring the full computational cost of retraining and extensive addtional simulations.
\gls{tl} offers a compelling solution to this problem by enabling the reuse of a pre-trained model on a new but related task~\citep{pan2009survey}.
In the context of \gls{dnn}s, \gls{tl} involves leveraging the knowledge acquired during training on one dataset to improve learning performance on a different dataset~\citep{yosinski2014transferable}.
This approach can substantially reduce the amount of training data needed, training time and computational resources while enhancing model generalization.
Instead of starting from the scratch, a new \gls{dnn} metamodel can be built on top of a pre-trained metamodel and fine-tuned on new data.
This allows the metamodel to adapt quickly to unseen conditions and new \gls{va} contracts.

In this chapter, we explore the application of \gls{tl} to the dynamic hedging of \gls{va} contracts using \gls{lstm} metamodels.
We propose a novel \gls{tl} framework that accelerates the training of \gls{dnn} metamodels for nested simulation procedures in dynamic hedging.
This setting is particularly relevant for dynamic hedging problems, as insurance companies often underwrite new contracts under changing market conditions and actuarial assumptions.
This chapter proposes a \gls{tl} framework that makes fast adaptation of metamodels possible.
Our approach involves pre-training an \gls{lstm} network on a large dataset of \gls{va} simulations and then fine-tuning it on a smaller dataset of new \gls{va} contracts.
We evaluate the performance of the \gls{tl} framework on a real-world dataset of \gls{va} contracts and compare it with training a \gls{lstm} metamodel from scratch.
More importantly, we show the strengths and weaknesses of different \gls{tl} techniques for rapid adaptation of \gls{lstm} metamodels.

The rest of this chapter is organized as follows.
Section~\ref{sec3:background} provides an overview of the dynamic hedging problem for \gls{va} contracts and the use of \gls{lstm} networks as metamodels in a nested simulation procedure.
Section~\ref{sec3:transfer_learning} introduces the \gls{tl} framework for rapid adaptation of \gls{lstm} metamodels in dynamic hedging.
Section~\ref{sec3:experiments} presents the experimental setup and results, comparing the performance of \gls{tl} with training from scratch.
Finally, Section~\ref{sec3:conclusion} concludes the chapter and discusses future research directions.

@@

\gls{tl} is a machine learning paradigm where knowledge acquired from a source task is utilized to improve learning performance on a related target task.
One of the primary applications of \gls{tl} in finance is asset price prediction. 
Traditional models, such as \gls{gbm}, \gls{arima} and \gls{garch}, have been widely used in time series modeling.
\gls{arima} models~\citep{box1970distribution} combine autoregression, differencing, and moving average components to capture linear relationships in time series data.
\gls{garch} models~\citep{bollerslev1990modelling}, extend \gls{arch} models to better capture volatility clustering, which is a key stylized fact of financial time series~\citep{cont2001empirical}.
However, these models often struggle to capture complex dynamics, e.g., regime changes~\citep{hamilton1989new} and extreme market events~\citep{embrechts2013modelling} that are prevalent in financial markets.
In Chapter~\ref{chap:project2}, our numerical experiment uses a \gls{rsgbm} as the underlying asset model.
For estimating tail risk measures, a metamodel-based nested simulation procedure is proposed to approximate the inner simulation.
We have shown \gls{dnn} metamodels like \gls{rnn} and \gls{lstm} networks have demonstrated substantial improvements in modeling temporal dependencies, and they save a substantial amount of simulation budget.
However, training these models from the scratch requires an extensive amount of simulation data, which may not always be available for specific assets or under certain market conditions.

@@

Regarding a nested simulation of \gls{va} contracts,~\cite{cheng2019fast} is the most relevant study to our work, where they proposed a \gls{tl} framework for fast valuation of large portfolios of \gls{va} contracts.
Instead of using stochastic kriging~\citep{gan2015valuation}, they employed a pre-trained \gls{dnn} to select the best representative scenarios from a large portfolio of \gls{va} contracts.
In our work, we focus on the application of \gls{tl} to accelerate the training of \gls{lstm} metamodels for a nested simulation in dynamic hedging of \gls{va} contracts.
The fine-tuned \gls{lstm} metamodels can be readily adapted to the two-stage procedure and the single-stage procedure in Chapter~\ref{chap:project2} to predict the contract losses under different scenarios.

In our context of \gls{dnn}s metamodeling-based simulation procedures for hedging \gls{va} contracts, \gls{tl} involves pre-training a \gls{dnn} metamodel where the simulation budget is abundant and then fine-tuning it on a smaller dataset of new contracts or market conditions.
Written on the same underlying asset, different \gls{va} contracts may share common features or exhibit similar patterns, especially temporal dependencies and regime changes in the underlying financial time series.
Similarly, two \gls{va} contracts with the same features but based on different underlying assets may have some shared characteristics that can be leveraged to improve the learning performance of the metamodel.
By transferring knowledge from a pre-trained model to a new but related task, \gls{tl} can substantially reduce the computational cost of training the metamodel on the target task.

@@

In this section, we propose a \gls{tl} framework for rapid adaptation of \gls{lstm} metamodels in dynamic hedging of \gls{va} contracts.
The goal is to leverage the knowledge acquired during training on a large dataset of \gls{va} simulations to improve the learning performance on a smaller dataset of new \gls{va} contracts.

@@

        \STATE \textbf{Multi-task learning:} define shared \gls{lstm} layers $\theta_0$ and task-specific fully connected layers $\theta_k$ for each task $k$. The \gls{lstm} layers are shared across all tasks $\{\mathcal{T}_k\}_{k=1}^K$.

@@

        \STATE \textbf{Layer freezing:} once the multi-task metamodel is trained, freeze the parameters of the shared \gls{lstm} layers $\theta_0$.

@@

In this section, we evaluate the performance of the \gls{tl} framework for rapid adaptation of \gls{lstm} metamodels in dynamic hedging of \gls{va} contracts.
The low noise dataset generated by the standard nested simulation procedure in Section~\ref{sec2:numerical} is used to train the source \gls{lstm} metamodels.
We consider \gls{tl} to two types of target tasks: 
\begin{itemize}
    \item the same contract but different underlying asset model, and
    \item different contracts but the same asset model.
\end{itemize}  
Similar to Section~\ref{sec2:numerical}, we use the standard nested simulation procedure to generate the dataset for the source and target tasks.
Serveral \gls{va} datasets generated with the standard nested simulation procedures under \gls{gbm} and \gls{rsgbm} asset models.
In this experiment, we use the same \gls{rsgbm} asset model as Chapter~\ref{chap:project2} to generate the dataset for the source and target tasks.
When following a static lapse model and a dynamic lapse model, the monthly lapse rate is given by Equation~\ref{eq:lapse_rate} and Equation~\ref{eq:dynamic_lapse_rate}, respectively.
The experiments are run on a machine with a AMD Ryzen 9 7900X processor with 32 GB of RAM and a Nvidia RTX $\num{3060}$ Ti GPU.

@@

Multi-task learning enables the \gls{lstm} metamodels to learn shared representations across related \gls{va} contracts.
In this section, we examine the performance of multi-task learning applied to two types of \gls{va} contracts, \gls{gmmb} and \gls{gmwb} with the dynamic lapse rates.
The simulation datasets contain $2,\!000$ samples for each contract type, and the \gls{lstm} metamodels are trained using multi-task learning. 
In our experiments, Algorithm~\ref{alg3:multiTaskLearning} is used to train the \gls{lstm} metamodels simultaneously to minimize the multi-task MSE loss function in Equation~\eqref{eq3:multiTaskLoss}.
We use individual task training as a baseline for comparison, where the \gls{lstm} metamodels are trained separately on the \gls{gmmb} and \gls{gmwb} contracts.
The objective is to assess how multi-task learning can improve the training efficiency and performance of both products compared to individual task training.

@@

Multi-task learning offers a robust framework for transferring knowledge across related simulation schemes in the \gls{lstm} metamodeling for nested simulations. 
By sharing representations through the \gls{lstm} layers, the model can learn more generalizable features that are beneficial across different \gls{va} contracts. 
The multi-task approach is particularly beneficial when training data for individual tasks is scarce, as it effectively pools information across tasks to enhance learning efficiency and predictive performance.
We have extended the idea of pooling in Chapter~\ref{chap:project1} and~\ref{chap:project2} to a higher level.
Furthermore, a metamodel trained with multi-task learning can serve as a pre-trained model. 
It is adaptable to various \gls{va} contracts and asset models, and it is a versatile tool for practical applications in dynamic hedging of \gls{va} contracts.
